DRAFT: v6.dev \\ Vibration Characterisation of Strollers and\\ Cargo Bicycles for Transporting Infants\\ \vspace{2mm} \large{Including Recommendations for Users, Designers, Manufacturers, and Researchers} \\

### 2025-09-20 main.tex
Version of paper text just after Jason reviewed entirety after last comments from co-authors before journal submission.

% Style Guide
% - use British English, exceptions: stroller vs buggy or pram, sidewalk vs pavement
% - active voice is preferred over passive voice, minimize passive
% - use past tense to describe what we did
% - try to avoid subsub sections (keep to two layers)
% - use siunitx for all units, surround with brackets [\si{\meter}], note the
%   common definitions, e.g. \si{\kph}
% - use tilde ~ to prevent line breaks in words that should stay together, e.g.
%   something~\cite{} or ISO~2631-1
% - only use "significant" when referring to statistical significance
% - use "infant", not "baby"
% - use "bicycle" instead of "bike"
% - use "amplitude spectrum/spectra" for what we plot
% - use "standard" not "norm"
% - use the oxford comma
% - TODO: acceleration vs accelerations ?
% - Data breakdown terms are "Session", "Trial" (of specific "Scenario"),
%   "Repetition"
% -  Referencing: Figure~\ref{fig:name}, Equation~\ref{eq:name},
%    Section~\ref{sec:name}, Appendix~\ref{app:name}, Table~\ref{tab:name}
% - inline urls \href{https://www.e.com}{my hyperlink}
% - do not use [h] for figures except in appendix, let latex place them where it
%   wants
% - figures should be 300 dpi and sized to the actual physical dimension in the
%   paper
% Other Notes
% - be careful about editing autogenerated tables, they will be overwritten

@@

\title{
  DRAFT: v6.dev \\
  Vibration Characterisation of Strollers and\\
  Cargo Bicycles for Transporting Infants\\
  \vspace{2mm}
  \large{Including Recommendations for Users, Designers, Manufacturers, and Researchers} \\
}
% Gabriele as first, Jason as last, alphabetical for intermediate (discuss in meeting if other method desired
\author{
Gabriele Dell'Orto \and
Brecht Daams \and
Riender Happee \and
Georgios Papaioannou \and
Arjo J. Loeve \and
Jesper Meijerink \and
Thomas Valk \and
Jason K. Moore\footnote{corresponding author: j.k.moore@tudelft.nl, +31 (0)15 278 3556}
}

@@

\color{red}
\section*{\centering Notice}
%
This draft report is not yet peer-reviewed and is subject to change with
revisions. These results are preliminary, so we do not recommend relying on them
until a supporting peer-reviewed paper is published.

\color{black}
\begin{abstract}
  Infants transported in strollers and cargo bicycles experience whole-body vibrations. Only little information was found on  these vibrations and the associated comfort and health effects on infants. To improve this, we measure the acceleration of  the seat pan of five strollers and two styles of cargo bicycles with dummy infants representing ages 0 to 9 months over six different road surfaces of varying road roughness, at typical travel speeds. The five strollers included three modern systems and two vintage systems, all with some level of suspension. 
One cargo bicycle had electric support that enabled speeds up to 25~\si{\kph}. We report average seat pan acceleration for 78 different scenarios and investigate the effects of road surface, vehicle, seat type, and travel speed. 
Using the ISO standard for whole-body vibration, we show that rough road surfaces, amplified by travel speed, may have health risks for travel durations as low as \SIrange{10}{30}{\min}. Vintage stroller designs from 50 years ago significantly reduce vibrations compared to  modern strollers tested, indicating the need to return to better suspensions. Cargo bicycles ridden over paver bricks at the maximum speed of an electric bicycle can cause accelerations that should likely be avoided, except for the shortest of durations, until more direct evidence of risks to infants is studied. Vibration standards are needed for infant transport testing and evaluation, designers need to incorporate better suspension for infants, and research is needed to develop clear and direct connections to health and comfort effects.
\end{abstract}

@@

\section{Introduction}
%Paragraph to introduce baby transport
When a baby is born, it is not able to walk and must
be either carried by adults or transported in suitable wheeled
transport. Until about a century ago,  babies mainly stayed at home, which is still customary in some non-Western cultures. A century ago, perambulators were
used to transport infants outdoors. However, mothers were advised that a playpen
in the garden is healthier than the use of perambulators, partly because vibrations generated by a perambulator were deemed not comfortable and even
harmful for babies~\cite{behrend1931}, even though in those days perambulators had ample suspension. Since about 1970, cars became ubiquitous, and the heavy and
sizeable perambulators became strollers: lighter and smaller products foldable for
easy transportation in cars. Over the years, new means of transportation came
into use for infants: child safety seats to use in cars, bicycles with child seats
and cargo bicycles with child seats or baby shells.

%Paragraph to introduce baby vibration in transportation
All means of transport cause vibration, which is transferred to the sitting or lying infant. Unlike the advice in the 1930s, vibration is seemingly
not a subject of attention in the present design of these products. On the contrary, modern strollers do not have much suspension, and (electric) cargo bicycles are increasingly used for infants with seats that offer marginal suspension or padding. We thoroughly reviewed the literature, but found very little about the maximal vibration load that babies and older children can receive during transport without discomfort or harm, and vibrations generated by infant transport are only sparsely investigated.
%Therefore the effect of transport on infant comfort and health can not easily be assessed. 
To establish the amount of vibration to which infants are subjected in
transportation products, we investigated the vibration that infants receive
during transport in strollers and cargo bicycles.

% Paragraph about motivation
Research on vibration comfort during cycling is a logical starting point for investigating similar road-induced vibrations for infants. Cycling is beneficial for
health~\cite{Oja1998} and greatly contributes to reducing CO\(_2\) emissions, congestion, and air pollution~\cite{Neves2019}. As a result, bicycles are
increasingly being used to deliver goods, for daily
commuting and leisure. Among all journeys in the EU, 20-40\% are by
bicycle or on foot, with bicycle trips most frequent in the Netherlands,
Denmark, and Sweden and less frequent in Finland. Various countries and employers provide incentives to increase the use of bicycles. However, the comfort experienced
by riders and passengers (i.e., infants/children in cargo bicycles), is 
a multifaceted ~\cite{ayachi2015identifying} challenge in their wide
acceptance. %Meanwhile, this is a topic that is of great interest and requires attention when it comes to infants' transportation with strollers or bicycles.
% Paragraph about comfort in bicycles
The comfort of bicycles~\cite{too1990biomechanics} is mainly affected by physical
comfort and the impact of environmental factors (weather, route geometry,
and road roughness).  
Physical comfort is related to mechanical (bicycle and component design),
biomechanical (whole-body vibrations, human body dynamics, and kinematics), and
physiological factors (individual characteristics, e.g. sex, body size, weight, etc.).
Although there is relevant literature exploring the impact of environmental
factors (i.e., irregular road surface quality, weather conditions) on cyclist
comfort~\cite{Stinson2003,Hagemeister2003,ayachi2015identifying,Verhoeven2017,Teixeira2020},
there is limited work on understanding the biomechanical and physiological
factors of comfort of bicycle passengers (infants/children in bicycle
seats, cargo bicycles, or trailers). A particular concern is the transport
of infants under one year of age, who cannot well express any experience of discomfort
or pain.
Similar concerns are raised about strollers, where infants are also
exposed to road-induced vibrations. 
In this paper, we will explore biomechanical comfort by assessing whole-body
vibrations experienced by infants transported in cargo
bicycles and strollers, and by evaluating potential health risks and discomfort.

% Paragraph about transmission of vibrations
Biomechanical comfort is mainly related to vibrations induced by road
irregularities and transmitted through the vehicle to the human body ~\cite{too1990biomechanics}. 
Road irregularities are a driving factor, as vibrations induced in children are not efficiently
isolated due to the lack of adequate suspension systems in most bicycle models~\cite{vasudevan2017comparison} and strollers.
The bicycle or stroller design and the seating configuration affect posture and vibration transmission~\cite{Brand2020,Verma2016}.
Children/infants can be
transported sitting or lying in a different location than the cyclist (e.g. above a wheel when in a bicycle seat) and in/on a different `seat'
(a car seat, bicycle seat, stroller seat, or cot). 
%These all influence the transmission of vibrations from the seat to the body and head.
Vibrations increase with travel speed, for example, when
running with a stroller or due to the increased power offered by electric
bicycles (with and without trailers).
For the latter, the average speed using electric bicycles and speed pedelecs is
19\% higher (21.0~\si{\kph} and up to 63\% higher (28.1~\si{\kph}),
respectively, than conventional bicycling~\cite{SWOV2022}.
%The intensity of these vibrations could become even more critical with the change of load, which could be related with age and size of the infant/children transported. 
However, there is limited literature exploring whole-body
vibrations of infants and children transported by bicycle trailers,
cargo bicycles, and strollers 
~\cite{Ota2012,Ota2014,Kanya-Forstner2020, vanDriessche2018,
Schwanitz2020,MalteChildComfort2022}, and literature focusses on comfort rather
than health risks.

% Paragraph about children in bicycles
One of the research fields closely related to the vibrations that act on infants
during transportation is that of inflicted head injury by shaking trauma
(IHI-ST). However, there is a lack of reliable, applicable, and validated injury
thresholds to assess IHI-ST risks~\cite{hutchinson2024modeling}. Similarly,
even if there are indications that vibration affects
comfort in bicycles or strollers, there are no standard vibration assessment methods or
guidelines available with respect to their design. ISO~2631~\cite{ISO2631}, the
international standard for assessing whole-body vibrations, was derived using
data sets on motion platforms, with adults as participants and sitting postures
as adopted in passenger vehicles rather than on bicycles.
Gao~et~al.~\cite{Gao2018} therefore derived vibration
comfort limits for adults riding bicycles, based on their
subjective feedback.
Furthermore, in ISO~2631-1 the exposure time to vibrations is not considered a factor
in the determination of comfort, despite having a clear impact on health risks
according to the European Directive (2002/44/EC, 2002)
\cite{directive2002directive}.

% Paragraph Potential effects of vibrations on adults
Despite the lack of knowledge and test methods with regard to comfort or inflicted head
injury by shaking trauma for children/infants on bicycles, there are clear
indications in the literature that these should be considered.
About 36\% of males and 42\% of females in a total sample of 900 cyclists had
various discomfort complaints~\cite{groenendijk1992sitting}.
Discomfort of both men and women occurred even during short bicycle trips
(\SIrange{3}{10}{\kilo\meter}). However, results from research into comfort of active
adult cyclists cannot be applied directly to infants or children sitting or
lying in transportation systems.

Schwanitz~et~al.~\cite{Schwanitz2020} tested a child bicycle
trailer on smooth asphalt, gravel, and cobblestones using child shaped sandbag dummies (baby and toddler sizes) using ISO~2631-1
weightings. Tyre pressure and number of
passengers had no significant effect on vibration magnitude, but road surface
and travel speed did. 
The baby-sized dummy (5~\si{\kilo\gram}) experienced 20\% higher vibration than
the toddler-sized dummy (10~\si{\kilo\gram}), while the measured values
generally exceeded the ISO~2631-1 limit for vibrational comfort of adults.
Rothhämel~et~al.~\cite{MalteChildComfort2022} tested a
trailer seat and a tadpole configuration three-wheel cargo bicycle on smooth
asphalt and cobblestones for speeds from \SIrange{10}{20}{\kph} using ISO~2631-1 weightings. They found a large
increase in vibration amplitude on cobblestones compared to asphalt in the
same speed range. The child trailer exhibited larger accelerations
than the cargo bicycle with the child seat between the front wheels. 
Child seats in trailers and cargo bicycles generally had
vibrations of higher magnitude than those experienced in a car seat travelling at 30~\si{\kph} on a rough road \cite{gromadowski2013analysis}. Rothhämel and
Liu~\cite{rothhamel2023comfort} conducted measurements with a
6.5~\si{\kilo\gram} mass representing an infant with various tyre pressures. All
their values fell into the ``extremely uncomfortable'' zone for ISO~2631-1.

In addition to studies with dummies or lumped masses representing
infants, Kanya-Forstner~\cite{Kanya-Forstner2020} measured acceleration in
bicycle child trailers with human subjects aged 12 months to 6 years over
asphalt and gravel terrain (average speed of 12~\si{\kph} over a
20~\si{\minute} ride). 
According to the health assessment using ISO~2631-1, the vibrations are similar
to those experienced in child seats in cars and illustrate moderate or low
health risk for a 2~\si{\hour} duration.
On the other hand, if not corrected for duration, the values point to moderate
and high risks. The type of road surface had the greatest influence on the levels of vibration
exposure, followed by type of trailer, while gel cushions as support did not
significantly influence the vibration measured at the seat/pan interface. Hence,
there is a need to refocus on the impact of vibrations on infants and
children's health and comfort~\cite{rybarczyk2020physiological} to ensure 
safe, comfortable and wide use of bicycles.

% Don't edit, working to change the following paragraph
In the context of infants transportation with stroller seats, Kok
Siong~\cite{lim2018study} carried out an indoor and an outdoor experiment to
capture the vibrations at the seat and backrest of a baby
stroller, using weights from \SIrange{4}{14.14}{\kilo\gram}, and  a child
subject of 10.30~\si{\kilo\gram} as well as a dummy weighing
10.30~\si{\kilo\gram}. 
Comfort levels (ISO~2631-1) for indoor testing were  `Fairly
uncomfortable' to `A little uncomfortable', where outdoor testing resulted in
`Extremely uncomfortable' and `Fairly uncomfortable' values. The child and dummy provided similar results. %However, the results still consider ISO~2631-1 standard, which is designed for adults and specific postures. 
Sierzputowski~
et~al.~\cite{sierzputowski2021pilot} tested a modern
and
an older stroller on asphalt road, concrete
paving blocks, concrete plates, dirt road, lawn and damaged concrete. Several conditions resulted in the highest discomfort levels ($>$2~\si{\mps\squared} according to ISO~2631-1). 
Okajima~et~al.~\cite{okajima2020dynamic} tested seven
children, sitting upright in a stroller, riding over a protrusion
(13x60 mm) with either one wheel or two wheels. 
The children were 3 to 6 years old, boys and girls, weighing
\SIrange{15.94}{19.96}{\kilo\gram} with a length of
\SIrange{98.0}{112.4}{\centi\meter}. 
The infant’s head and chest exhibited strong vibrations at
1~\si{\hertz} along all three axes (x, y and z), vibration along the
y-direction at 2~\si{\hertz}, with limited vibration above
8~\si{\hertz}. 
%This result is critical as this range of frequencies is very common in bicycles and strollers. 
However, they did not assess comfort and health risk.

The above studies provide experimental evidence of strong vibrations in real children, dummies, or simple masses transported in cargo bicycles, bike trailers, and strollers.
Most studies use ISO~2631-1 frequency weighting and report very or extremely uncomfortable conditions with higher speeds and rough surfaces, while only one study evaluated health effects. Only a few studies focus on the effect of vibration on babies and infants under 12 months of age.

This paper addresses both comfort and potential health effects for babies and infants.
We equipped five
strollers and two cargo bicycles with inertial measurement units (IMU) and
carried out 78 experiments on different road surfaces and with different travel speeds, using
dummies representing 0, 3, and 9 month-old infants.
We present results using ISO~2631-1
frequency weightings. 
Given that ISO-2631-1 is neither validated for children nor for lying postures, we also report unweighted results to assess power spectrum and bandwidth.
We investigated the influence of vehicle, seat, body size, speed, and road surface
and provide recommendations to users, designers, and future researchers.

\section{Methods}
%
\subsection{Test Equipment}
%
Vibration measurements were conducted at approximately constant speed using five
different strollers and two cargo bicycles with two baby seats, chosen for their
popularity and/or distinctive features. The strollers consisted of three modern
strollers: Maxi-Cosi Street Plus (Dorel Juvenile, Foxborough, USA), BABYZEN YOYO
0+ (Stokke, Ålesund, Norway), Fox 5 (Bugaboo, Amsterdam, The Netherlands) and
two vintage strollers from unknown manufacturers ``Green Machine'' cot-style and
``Old Rusty'' seat-style. The cargo bicycles were a two-wheel UrbanArrow (Pon,
Amsterdam, The Netherlands) and a delta configuration tricycle (Keiler Bakfiets, The Netherlands) which were each fitted with the two seats: Joolz Pebble Pro (Dorel Juvenile, Foxborough, USA), Melia Baby Shell (Melia, Rotterdam, The Netherlands).
~\footnote{Appendix~\ref{app:equipment} provides.
figures and technical descriptions of all strollers and cargo bicycles.}

The tests involved three baby dummies with sizes and weights representative of
newborns and infants 3 and 9 months of age. Newborns (`0 months') are the
youngest (smallest, lightest) children that are transported in strollers with a
cot. Three-month-old infants are the youngest (smallest, lightest) children that
are transported in cargo bicycles, fixed in a baby car seat or baby shell.
Nine month old infants are the youngest (smallest, lightest) that can sit up
straight by themselves and thus can be transported sitting up straight, in a
stroller with a seat or on the bench of a cargo bicycle.

We considered the use of ``crash test dummies'' (also known as anthropomorphic
test devices or ATDs), which are designed and commonly used to test child seats
in car crash tests. However, the closest available body sizes for crash test
dummies do not meet the body sizes desired for this investigation. Furthermore, crash test dummies are of a fundamentally
different design, based on scaling of adult biomechanical data
rather than child data and with dynamic properties designed for crash
conditions, not for vibration testing. Therefore crash test dummies were not considered a good option to use for the envisaged vibration testing.

Testing with real infants would be unethical for the most severe conditions and create challenges in
terms of reproducibility. Hence we bought and adapted ``reborn'' dummies
available in a wide range of postures from a webshop specialising in dolls and
dummies (\href{http://www.atelier-wiesje.nl}{Atelier Wiesje}). The dummies were
filled with a mixture of cat litter, sand, and water to reproduce the exact
weights needed. Cat litter sand features smooth and round shapes
grains, which do not damage the dummies' materials (fabric and plastic). The
weight and body height of the dummies were made to match the average measures
for children of 0, 3 and 9 months old: 3.48~\si{\kg}|50~\si{\cm}
5.90~\si{\kg}|62.5~\si{\cm}, and 8.90~\si{\kg}|70~\si{\cm}, respectively.

%We wanted to use the measurements from Steenbekkers~\cite{Steenbekkers1993}. However, this research provided the results only per three-month group, which did not fit our age categories. 
We used Dutch infant growth charts to determine the average weight and body
height at age 0 months [reference needed!] and French growth charts for 3 and 9
months. French children are the smallest in Europe, and Dutch children are the
tallest. At 3 months, average Dutch children weigh 113 grams more than average
French children and at 9 months, the difference is 85 grams.

>>>>

% Style Guide
% - use British English, exceptions: stroller vs buggy or pram, sidewalk vs pavement
% - active voice is preferred over passive voice, minimize passive
% - use past tense to describe what we did
% - try to avoid subsub sections (keep to two layers)
% - use siunitx for all units, surround with brackets [\si{\meter}], note the
%   common definitions, e.g. \si{\kph}
% - use tilde ~ to prevent line breaks in words that should stay together, e.g.
%   something~\cite{} or ISO~2631-1
% - only use "significant" when referring to statistical significance
% - use "infant", not "baby"
% - use "bicycle" instead of "bike"
% - use "amplitude spectrum/spectra" for what we plot
% - use "standard" not "norm"
% - use the oxford comma
% - TODO: acceleration vs accelerations?
% - Data breakdown terms are "Session", "Trial" (of specific "Scenario"),
%   "Repetition"
% -  Referencing: Figure~\ref{fig:name}, Equation~\ref{eq:name},
%    Section~\ref{sec:name}, Appendix~\ref{app:name}, Table~\ref{tab:name}
% - inline urls \href{https://www.e.com}{my hyperlink}
% - do not use [h] for figures except in appendix, let latex place them where it
%   wants
% - figures should be 300 dpi and sized to the actual physical dimension in the
%   paper
% Other Notes
% - be careful about editing autogenerated tables, they will be overwritten

@@

\title{
  DRAFT: v6.dev \\
  Vibration Characterisation of Strollers and\\
  Cargo Bicycles for Transporting Infants\\
  \vspace{2mm}
  \large{Including Recommendations for Users, Designers, Manufacturers, and Researchers} \\
}

\author{
Gabriele Dell'Orto \and
Brecht Daams \and
Riender Happee \and
Georgios Papaioannou \and
Arjo J. Loeve \and
Jesper Meijerink \and
Thomas Valk \and
Jason K. Moore\footnote{corresponding author: j.k.moore@tudelft.nl, +31 (0)15 278 3556}
}

@@

\color{red}
\section*{\centering Notice}
%
\begin{center}
  This paper has been submitted for peer review and is subject to change with
  revisions.
\end{center}

\color{black}
\begin{abstract}
  % NOTE: keep to 150 words max
  We evaluated road-induced vibrations with dummies representing infants aged 0,
  3, and 9 months with five strollers and two cargo bicycles on six common road
  surfaces using the ISO standard for whole-body vibration. Strollers induced on
  average 0.4~\si{\mps\squared} on tarmac and up to 5.0~\si{\mps\squared} on
  cobblestones at a mean walking speed of 5.3~\si{\kph}. Cargo bicycles induced
  induced as low as 0.6~\si{\mps\squared} on tarmac and up to
  10.7~\si{\mps\squared} at 25~\si{\kph} on paver bricks. The standard suggests
  that the highest accelerations for the strollers and cargo bicycles should be
  limited to continuous duration less than 10~\si{\minute}. Effects of body size
  and posture were minor. Two vintage stroller designs significantly reduced
  vibrations compared to three modern strollers, indicating the benefits of
  compliant suspensions. 
\end{abstract}

\section*{Keywords}
%
whole-body vibration, infant, bicycle, stroller, ISO

@@

\section{Introduction}
% Paragraph to introduce baby transport
When a baby is born, it is not able to walk and must be carried by others or
transported in a suitable wheeled transport. Until about a century ago, babies
were mainly kept at home, which is still customary in some non-western cultures.
More than a century ago, perambulators were introduced to transport infants
outdoors. However, mothers were advised that a playpen in the garden is
healthier than the use of perambulators, in part because vibrations generated by
a perambulator were deemed not comfortable and even harmful to
babies~\cite{behrend1931}, even though in those days perambulators had ample
suspension. Since about 1970, cars became ubiquitous and the heavy and sizeable
perambulators became strollers: lighter and smaller products foldable for easy
transportation in cars. Over the years, new means of transportation came into
use for infants: child safety seats to use in cars, bicycles with child seats,
and cargo bicycles with child seats or baby shells.

%Paragraph to introduce baby vibration in transportation
All means of transport cause vibration, which is transferred to the sitting or
lying infant. Unlike the advice in the 1930's, vibration is seemingly not a
subject of attention in the present design of these products. On the contrary;
modern strollers do not have much suspension, and (electric) cargo bicycles are
increasingly used for infants with seats that offer marginal suspension or
padding. We thoroughly reviewed the literature, but found very little about the
maximal vibration load that babies and older children can receive during
transport without discomfort or harm, and vibrations generated by infant
transport are only sparsely investigated. To establish the amount of vibration
to which infants are subjected in transportation products, we measured seat pan
vibrations while transporting infant dummies in strollers and cargo bicycles
with a focus on potential health risks and discomfort.

% Paragraph about motivation
Research on vibration comfort during cycling is a logical starting point for
investigating similar road-induced vibrations for infants. Cycling is beneficial
for health~\cite{Oja1998} and greatly contributes to reducing CO\(_2\)
emissions, congestion, and air pollution~\cite{Neves2019}. As a result, bicycles
are increasingly being used to deliver goods, for daily commuting and for
leisure. Among all journeys in the EU, 20-40\% are by bicycle or on foot, with
bicycle trips most frequent in the Netherlands, Denmark, and Sweden and least
frequent in, for example, Finland. Various countries and employers offer
incentives to increase bicycle use. However, the comfort experienced by riders
and passengers (e.g. infants/children in cargo bicycles), is a
multifaceted~\cite{ayachi2015identifying} challenge to wide acceptance of
bicycling.

% Paragraph about comfort in bicycles
The comfort of bicycles~\cite{too1990biomechanics} is mainly affected by
physical comfort and the impact of environmental factors (weather, route
geometry, and road roughness). Physical comfort is related to mechanical
(bicycle and component design), biomechanical (whole-body vibrations, human body
dynamics, and kinematics), and physiological factors (individual
characteristics, e.g. sex, body size, weight). Although there is relevant
literature exploring the impact of environmental factors (i.e., irregular road
surface quality, weather conditions) on cyclist
comfort~\cite{Stinson2003,Hagemeister2003,ayachi2015identifying,Verhoeven2017,Teixeira2020},
there is only limited work on understanding the biomechanical and physiological
factors of comfort of bicycle passengers (infants/children in bicycle seats,
cargo bicycles, or trailers). A particular concern is the transport of infants
under one year of age, who cannot well express any experience of discomfort or
pain. Similar concerns are raised about strollers, where infants are also
exposed to road-induced vibrations.

% Paragraph about transmission of vibrations
Biomechanical comfort is mainly related to vibrations induced by road
irregularities and transmitted through the vehicle to the human
body~\cite{too1990biomechanics}. Road irregularities are a driving factor, as
vibrations induced in children are not efficiently isolated due to the lack of
adequate suspension systems in most bicycle
models~\cite{vasudevan2017comparison} and strollers. The bicycle or stroller
design and the seating configuration affect posture and vibration
transmission~\cite{Brand2020,Verma2016}. Children/infants can be transported
sitting or lying in a different location than the cyclist (e.g. above a wheel
when in a bicycle seat) and in/on a different `seat' (a car seat, bicycle seat,
stroller seat, or cot). Vibrations generally increase with travel speed, for
example, when running with a stroller or due to the increased power offered by
electric bicycles (with and without trailers). For the latter, the average speed
using electric bicycles and speed pedelecs is 19\% higher (21.0~\si{\kph} and up
to 63\% higher (28.1~\si{\kph}), respectively, than conventional
bicycling~\cite{SWOV2022}. The intensity of these vibrations could become even
more critical with the change of load, which is related with age and size of the
infant/children transported. However, there is limited literature exploring
whole-body vibrations of infants and children transported by bicycle trailers,
cargo bicycles, and strollers~\cite{Ota2012,Ota2014,Kanya-Forstner2020,
vanDriessche2018, Schwanitz2020,MalteChildComfort2022}. Most literature focusses
on comfort rather than health risks.

% Paragraph about children in bicycles
One of the research fields closely related to the vibrations that act on infants
during transportation is that of inflicted head injury by shaking trauma
(IHI-ST). However, there is a lack of reliable, applicable, and validated injury
thresholds to assess IHI-ST risks~\cite{hutchinson2024modeling}. Similarly, even
if there are indications that vibration affects comfort in bicycles or
strollers, there are no standard vibration assessment methods or guidelines
available with respect to their design. ISO~2631~\cite{ISO2631}, the
international standard for assessing whole-body vibrations, was derived using
data sets on motion platforms, with adults as participants and sitting postures
adopted in passenger vehicles rather than on bicycles. Gao~et~al.~\cite{Gao2018}
show this discrepancy with vibration comfort limits for adults riding bicycles,
based on their subjective feedback, tolerating more than the the vibration
limits suggested by ISO~2631-1. Furthermore, in ISO~2631-1 the exposure time to
vibrations is not considered a factor in the determination of comfort, despite
having a clear impact on health risks according to the European Directive
(2002/44/EC, 2002)~\cite{directive2002directive}.

% Paragraph Potential effects of vibrations on adults
Despite the lack of knowledge and test methods with regard to comfort or
inflicted head injury by shaking trauma for children/infants on bicycles, there
are clear indications in the literature that these should be considered. About
36\% of males and 42\% of females in a total sample of 900 cyclists had various
discomfort complaints~\cite{groenendijk1992sitting}. Discomfort occurred even
during short bicycle trips (\SIrange{3}{10}{\kilo\meter}). However, the
results of research on the comfort of active adult cyclists cannot be applied
directly to infants or children sitting or lying in transportation systems.

Schwanitz~et~al.~\cite{Schwanitz2020} tested a child bicycle trailer on smooth
tarmac, gravel, and cobblestones using child shaped sandbag dummies (baby and
toddler sizes) using ISO~2631-1 weightings. Tyre pressure and number of
passengers had no significant effect on vibration magnitude, but the road
surface and travel speed did. The infant dummy (5~\si{\kilo\gram}) experienced a
20\% higher vibration than the toddler dummy (10~\si{\kilo\gram}), while the
measured values generally exceeded the ISO~2631-1 limit for vibrational comfort
of adults. Rothhämel~et~al.~\cite{MalteChildComfort2022} tested a trailer seat
and a tadpole configuration three-wheel cargo bicycle on smooth tarmac and
cobblestones for speeds from \SIrange{10}{20}{\kph} using ISO~2631-1 assessment.
They found a large increase in vibration amplitude on cobblestones compared to
tarmac in the same speed range. The child trailer exhibited larger
accelerations than the cargo bicycle with the child seat between the front
wheels. Child seats in trailers and cargo bicycles generally had vibrations of
higher magnitude than those experienced in a car seat travelling at 30~\si{\kph}
on a rough road \cite{gromadowski2013analysis}. Rothhämel and
Liu~\cite{rothhamel2023comfort} also conducted laboratory measurements with a
6.5~\si{\kilo\gram} mass representing an infant in a bicycle trailer with
various tyre pressures. All their values fell into the ``extremely
uncomfortable'' zone for ISO~2631-1.

In addition to studies with dummies or lumped masses representing infants,
Kanya-Forstner~\cite{Kanya-Forstner2020} measured acceleration in bicycle child
trailers with human subjects aged 12 months to 6 years over tarmac and gravel
terrain (average speed of 12~\si{\kph} over a 20~\si{\minute} ride). According
to the health assessment using ISO~2631-1, the vibrations are similar to those
experienced in child seats in cars and illustrate moderate or low health risk
for a 2~\si{\hour} duration. She points out the importance of correcting for
duration, otherwise the values point to moderate and high risks. The type of
road surface had the greatest influence on the levels of vibration exposure,
followed by the type of trailer, while the gel cushions as support did not
significantly influence the vibration measured at the seat/pan interface. Hence,
there is a need to refocus on the impact of vibrations on infants and children's
health and comfort~\cite{rybarczyk2020physiological} to ensure safe, comfortable
and widespread use of bicycles.

In the context of infant transportation with stroller seats, Kok
Siong~\cite{lim2018study} carried out an indoor and an outdoor experiment to
capture the vibrations at the seat and backrest of a baby stroller, using
weights from \SIrange{4}{14}{\kilo\gram}, and a child subject of
10.30~\si{\kilo\gram} as well as a dummy weighing 10.30~\si{\kilo\gram}. Comfort
levels (ISO~2631-1) for indoor testing were  `Fairly uncomfortable' to `A little
uncomfortable', where outdoor testing resulted in `Extremely uncomfortable' and
`Fairly uncomfortable' values. The child and dummy provided similar results.
Sierzputowski~et~al.~\cite{sierzputowski2021pilot} tested a modern and an older
stroller on tarmac road, concrete paving blocks, concrete plates, dirt road,
lawn, and damaged concrete. Several conditions resulted in the highest
discomfort levels ($>$2~\si{\mps\squared} according to ISO~2631-1).
Okajima~et~al.~\cite{okajima2020dynamic} tested seven children, sitting upright
in a stroller, riding over a protrusion (13$\times$60~\si{\milli\meter}) with
either one wheel or two wheels. The children were 3 to 6 years old, boys and
girls, weighing \SIrange{15.94}{19.96}{\kilo\gram} with a length of
\SIrange{98.0}{112.4}{\centi\meter}. The heads and chests of the children
exhibited strong vibrations at 1~\si{\hertz} along all three axes (x, y, and z),
vibration along the y-direction at 2~\si{\hertz}, with limited vibration above
8~\si{\hertz}.

The above studies provide experimental evidence of potentially worrying
vibrations in real children, dummies, or simple masses transported in cargo
bicycles, bike trailers, and strollers. Most studies use ISO~2631-1 frequency
weighting and report very or extremely uncomfortable conditions with higher
speeds and rough surfaces, while only one study evaluated health effects. Only a
few studies focus on vibrations experienced by infants under 12 months of age.

This paper addresses both comfort and potential health effects for infants. We
equipped five strollers and two cargo bicycles with inertial measurement units
(IMU) and carried out 67 experiments on different road surfaces and with
different travel speeds, using dummies representing 0, 3, and 9-month-old
infants. We assessed the data and derived results using ISO~2631-1 frequency
weightings and procedures. Given that ISO-2631-1 is neither validated for
children nor for lying postures, we also report unweighted results to assess
power spectrum and bandwidth. We investigated the influence of vehicle, seat,
body size, speed, and road surface and provided recommendations to users,
designers, and for future research.

\section{Materials and Methods}
%
\subsection{Test Equipment}
%
Vibration measurements were conducted using five different strollers and two
cargo bicycles with two baby seats, chosen for their popularity and/or
distinctive features. The strollers included three modern strollers:
\underline{Bugaboo} Fox 5 (Bugaboo, Amsterdam, The Netherlands),
\underline{Maxi-Cosi} Street Plus (Dorel Juvenile, Foxborough, USA), and
\underline{Stokke} BABYZEN YOYO 0+ (Stokke, Ålesund, Norway) and two vintage
strollers from unknown manufacturers ``\underline{Green Machine}'' cot-style and
``\underline{Old Rusty}'' seat-style. The cargo bicycles were a delta
configuration tricycle \underline{Keiler} (Keiler Bakfiets, The Netherlands) and
a two-wheel \underline{Urban Arrow} (Pon Bicycle Holding B.V., Amsterdam, The
Netherlands). Each was tested with both seats: Maxi-Cosi X Joolz
\underline{Pebble} Pro i-Size (Dorel Juvenile, Foxborough, USA) and
\underline{Melia} Baby Shell (Melia, Rotterdam, The Netherlands). The underlined
words for each product designate an abbreviated name used in figures and tables
in the paper and Figure~\ref{fig:equipment} shows the tested
products.~\footnote{Appendix~\ref{app:equipment} provides more detailed figures
and technical descriptions of all tested strollers and cargo bicycles.}
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{fig/equipment.png}
  \caption{Tested products: cargo bicycles with baby seats, strollers, and
  dummies. Images are reproduced from manufacturer's websites except for the
  vintage strollers and dummies.}
  \label{fig:equipment}
\end{figure}

The tests involved three baby dummies with sizes and weights representative of
infants 0, 3 and 9 months of age. Newborns (`0 months') are the youngest
(smallest, lightest) children that are transported in strollers with a cot.
3-month-old infants are the youngest (smallest, lightest) children that are
transported in cargo bicycles, fixed in a baby car seat or baby shell. At nine
months on average, infants can sit up straight by themselves and thus can be
transported sitting up straight, in a stroller with a seat or on the bench of a
cargo bicycle. Testing with real infants would be unethical for the most severe
conditions and create challenges in terms of reproducibility. Hence we bought
and adapted commercially available ``Reborn'' dummies
(\href{http://www.atelier-wiesje.nl}{Atelier Wiesje}), see
Figure~\ref{fig:equipment}~where Appendix  {\ref{app:equipment} provides full
details}.

The weight and body height of the dummies were made to match the average
measures for children of 0, 3 and 9 months old: 3.48~\si{\kg}|50~\si{\cm}
5.90~\si{\kg}|62.5~\si{\cm}, and 8.90~\si{\kg}|70~\si{\cm}, respectively. We
used Dutch infant growth charts to determine the average weight and body height
at age 0 months~\cite{TNO/LUMC1998} and French growth charts for 3 and 9
months~\cite{FrenchGrowthChart} as requested by the funder; with French children
being the smallest in Europe, while the Dutch are the tallest . The dummies were
filled with a mixture of cat litter, sand, and water to reproduce the mass and
mass distribution according to the body segment information reported
in~\cite{Snyder1977}.
